Untitled

### 2022-03-08 Manuscript/Main_text/Title_page_Oikos.Rmd
Edits on text

---
title: The non-random assembly of functional motifs in plant-pollinator networks worldwide
editor_options:
  chunk_output_type: console
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{caption,setspace}
- \captionsetup[figure]{labelformat=empty}
- \floatplacement{figure}{H}
- \usepackage[skip=3pt]{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{lscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output:
  pdf_document: default
  html_document: default
  word_document: default
always_allow_html: true
bibliography: bibliography.bib
---

__Jose B. Lanuza^1,2^__  | __Alfonso Allen-Perkins^2,3^__ |  __Romina Rader^1^__ | __Ignasi Bartomeus^2^__

@@

\captionsetup[table]{labelformat=empty}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

\singlespacing
{\normalfont Ecological processes leave distinct structural imprints on species interactions shaping the topology of mutualistic networks. Detecting those relationships is not trivial since they go beyond pair-wise interactions, but may get blurred when considering full network descriptors. However, recent work has shown the network meso-scale can capture this important information. The meso-scale describes network subgraphs representing patterns of interactions between a small number of species (i.e. motifs) and those constitute the building blocks of the whole network. Here, we have compiled 60 networks from 18 different studies and show that some motifs are consistently under- and over-represented worldwide, suggesting that the building blocks of plant-pollinator networks are not random. Furthermore, we found that distinct motif positions describing species ecological roles (e.g., generalization, indirect interactions) are occupied by different functional groups on both trophic levels. For instance, bees appear consistently in generalist motif positions and plants with higher levels of outcrossing tend to have a large number of indirect interactions. Interestingly, the realized combinations of different functional groups within a motif can not be retrieved from their joint probability distributions, indicating that functional group combinations are not random either. Our result highlight the non-random structure of indirect interactions in plant-pollinator networks and opens the door to predict novel interactions beyond pairwise interactions within a community context.}

\doublespacing

# INTRODUCTION

Ecological communities are formed by a plethora of interacting species that form networks of interactions. Because of the tremendous complexity of these ecological networks, species interactions are generally condensed in metrics that summarize this information [@guimaraes2020]. Plant-pollinator interactions are no exception, and they are often studied with indices that aggregate the information at the species (micro-scale) or network level (macro-scale). For example, plant-pollinator network level approaches have identified common invariant structural properties across networks, including a degree distribution that decays as a power law [@jordano1987], nestedness [@bascompte2003], or modularity [@olesen2007]. In addition, the local species position within the network can define its degree of specialization [@bluthgen2006] or its role connecting the rest of the community [@olesen2007]. Despite the unquestionable progress with the use of these metrics at both species and network level, condensing the information into a single metric implies the loss of relevant ecological information that obscure the understanding of the ecological processes [@cirtwill2018; @simmons2019].

Traditionally, plant-pollinator research has focused on direct interactions but overlooked indirect interactions (i.e., the mediated effect between two species by a third species), such as facilitative or competitive interactions between plants for pollinators [@moeller2004; @sargent2008]. Despite the widespread nature of indirect interactions in ecological communities [@strauss1991], plant-pollinator research often fails to finely capture those indirect interactions with the conventional analytical tools that condense the information either by species (e.g., interaction frequency) or in global topological indices (e.g., nestedness). Nonetheless, the emerging framework of network motifs (small sub-networks recorded within a given network) in plant-pollinator research allows to consider both direct and indirect interactions [@simmons2019]. Motifs are often referred to as the building blocks of a network that depict subsets of interactions [@milo2002]. The analysis of network motifs in plant-pollinator networks have revealed that the different ecological processes that govern species interactions (e.g., species abundances versus trait-matching) can lead to different patterns of indirect interactions [@simmons2020]. Yet, the global patterns of indirect interactions in real plant-pollinator networks are still unknown (e.g., over- and under- represented motifs).

Despite motifs are abstract representations often decoupled from species ecology, linking the structural properties of the meso-scale with the species ecology can help the understanding of ecological processes. For instance, different motifs can have different ecological meanings [@simmons2019] and the position within a motif can determine the species functional role [@stouffer2012; @baker2015]. However, it is unclear how the species ecology and life history traits determines the species' functional role within the network of interactions [@coux2016]. For example, large pollinators can forage larger distances [@greenleaf2007], deposit greater pollen quantities [@foldesi2021] and handle complex zygomorphic flowers in comparison with small pollinators that are restricted to lower floral complexity [@gong2009]. How this different pollinator behavior translates into their interaction topology is unknown. Similarly, recent empirical findings indicate that the meso-scale is the best descriptor of plant reproductive success [@allen2021], but little is known about how plants reproductive strategies shape their position within the network of interactions. Although some studies have evaluated plant reproductive strategies in plant-pollinator networks [@tur2013; @lazaro2020], this is an often overlooked aspect in a community context [@devaux2014] and rarely incorporated into plant-pollinator network studies. Hence, exploring how the main plant reproductive strategies integrate with the emergent motif framework can shed light on key aspects of ecosystem functioning.

Here, we used 60 plant-pollinator networks from 18 different studies and 14 countries, alongside functional groups of plants and floral visitors based on functional traits and taxonomic rank, respectively. To obtain plant functional groups, we used a comprehensive dataset that included floral, reproductive and vegetative traits compiled in a larger set of networks. Floral visitor functional groups were divided by the main taxonomic groups that differed in life form and behaviour. Once we split the different plant-pollinator networks into their motif elements, we explored: (i) if there is a common invariant structural property in the overall motif networks (i.e. over and under-represented motifs); (ii) which functional groups are over and under-represented in the different motif positions; and, (iii) if there are over and under-reprsented functional group combinations within a motif.

# METHODS

**Plant-pollinator studies**

We have compiled 60 plant-pollinator networks from 18 different studies (**Table S1**). All studies sampled plant-pollinator interactions in natural systems and were selected based on wide geographical coverage and presence of interaction frequency as a measure of interaction strength. In total, there were 503 plant species, 1,111 floral visitors species and 6,248 pairwise interactions registered. For ease of data manipulation, plant and floral visitors species names were standardize with the help of the R package _taxize_ version _0.9.99_ [@chamberlain2020]. All analyses and data manipulation were conducted in R _version 4.0.5_ [@R].

**Plant and floral visitors functional groups**

First, plant species were grouped through hierarchical clustering into the optimal number of functional groups that summarized the main plant reproductive strategies. For this, we used the trait dataset collated in Chapter 1 that comprised a total of 1506 species including the 503 species considered in this study (Supplementary material Chapter 1 **Table S1**). This dataset consisted on 8 floral, 4 reproductive and 3 vegetative traits excluding traits with high percentage of missing values (over 30$\%$; **Table S2**). We opted to calculate the plant functional groups on this larger set of species because of the higher accuracy when delimiting functional groups with that many variables and species [@dolnicar2014]. To feed the clustering analysis, we calculated the distance between the different qualitative and scaled quantitative variables with Gower distance [@gower1971]. For this, we used the function _gowdis_ with method _ward.D2_ from the package _FD_ version _1.0-12_ [@laliberte2014]. Finally, we conducted hierarchical clustering with the function _hclust_ from the _R_ stats package version _4.0.5_ and calculated the optimal number of clusters with the function _kgs_ from the package _maptree_ version _1.4-7_ [@white2009].

Second, floral visitors were grouped into functional groups based on taxonomic rank as done similarly in past plant-pollinator studies [@fenster2004; @ollerton2009]. We opted to divide floral visitors on the taxonomic rank level and not with functional traits because the main orders of floral visitors differed in form and behaviour and had lower superior taxonomical complexity (i.e., floral visitors had 6 orders versus plants that had 38). Thus, this allowed us to group floral visitors into functional groups that represented adequately the main life strategies of the possible pollinators: (i) bees (Hymenoptera-Anthophila), (ii) non-bee Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Syrphidae-Diptera), (iv) non-syrphids-Diptera, (v) Lepidoptera and (vi) Coleoptera. However, a minor set of species belonged to other functional groups that were considered in analyses but not discussed further because of their low representation in the full set of networks (3.55$\%$ of the total interactions recorded). These functional groups were 'lizards', 'birds' and 'other insects'. This last group was formed by a mix of uncommon insect taxa on the full set of networks.

**Overall motif patterns**

Following previous work [@simmons2019; @simmons2020], we broke down the plant-pollinator networks into their constituent motifs. Prior to analyses, we turned the quantitative networks into qualitative (or binary) ones, where interactions are present or absent. To test the effect of singletones, we ran all analysis to the same set of networks, but considering only interactions whose frequency was greater than one (64.98$\%$ of interactions). Results are qualitatively consistent and are not further discussed in the main text (**Figures S1 and S2**).

We calculated the frequency of all motifs up to five nodes (see **Figure 1**) for each empirical network, by using the _bmotif_ package version _2.0.2_ [@bmotifsimmons2019]. To control for variation in network size, motif frequencies were normalised as a proportion of the total number of motifs within each motif class (i.e., the number of nodes a motif contains). Like @simmons2020, we just used five-node motifs in our analyses for visualisation, interpretation and computational reasons. In addition, we also excluded two-node motifs (or links) from our analyses because their normalised frequencies would always equal one.

```{r, echo=FALSE, message=FALSE, cache=FALSE, fig.pos="t", warning=FALSE,out.width="75%",fig.align="center", fig.cap="\\textbf{Figure 1.} Adapted figure of Simmons et al., 2019 with all the possible motifs from two to five species in bipartite networks. There is a total of 17 possible motifs with 46 different positions denoted within each node.", fig.width=10,fig.height=8}
library(tidyverse)
library(RColorBrewer)
source("../Scripts/Scripts_Alfonso/plot_motif_positionsV2.R")

motif_number <- 1
p1 <- plot_motif_positions(motif_number)

motif_number <- 2
p2 <- plot_motif_positions(motif_number)

motif_number <- 3
p3 <- plot_motif_positions(motif_number)

motif_number <- 4
p4 <- plot_motif_positions(motif_number)

motif_number <- 5
p5 <- plot_motif_positions(motif_number)

motif_number <- 6
p6 <- plot_motif_positions(motif_number)

motif_number <- 7
p7 <- plot_motif_positions(motif_number)

motif_number <- 8   
p8 <- plot_motif_positions(motif_number)

motif_number <- 9
p9 <- plot_motif_positions(motif_number)

motif_number <- 10
p10 <- plot_motif_positions(motif_number)

motif_number <- 11
p11 <- plot_motif_positions(motif_number)

motif_number <- 12
p12 <- plot_motif_positions(motif_number)

motif_number <- 13
p13 <- plot_motif_positions(motif_number)

motif_number <- 14
p14 <- plot_motif_positions(motif_number)

motif_number <- 15
p15 <- plot_motif_positions(motif_number)

motif_number <- 16
p16 <- plot_motif_positions(motif_number)

motif_number <- 17
p17 <- plot_motif_positions(motif_number)

library(patchwork)

(p1 & ylab("2 species")| p2 & ylab("3 species") | p3 )/ (p4 & ylab("4 species")| p5 | p6 | p7)/ (p8  & ylab("5 species")| p9 | p10 | p11 | p12) / (p13 & ylab("5 species")| p14 | p15 | p16 | p17) & theme(axis.title.y = element_text(color="black", size=14, face="bold")) #+
 #plot_annotation(tag_levels = '1')

```

To assess the significance of the observed frequencies, we created 1,000 simulated networks for each binary network using the _nullmodel_ function and the _vaznull_ model in the _bipartite_ package version _2.14_ [@dormann2009]. Generated networks had the same number of plants and floral visitors, as well as the same connectance of their corresponding empirical networks. After extracting the motif frequencies from the simulated networks, for each motif type and empirical network, we calculated the percentage of simulated networks whose frequencies were smaller than the ones observed, that is, we estimated the percentile of the observed motif frequencies. Motifs whose percentile is close to 0 or 100 are under- or over-represented in the empirical networks, respectively, and thus they cannot be predicted by connectance and the number of species alone. To summarize general patterns across networks, we used an intercept-only linear mixed model (LMM) per motif with the help of the package lmer version _1.1-21_ [@bates2015], where the response variable was the observed motif percentile per network. In these models, we used the study identifiers in **Table S1** as a random intercept. By doing so, we obtained estimates of the average motif frequency, but controling the variation at the study level.

**Over and under-represented functional groups on motif positions**

We calculated which functional groups were over or under-represented in different motif positions by comparing position frequencies of empirical networks with those of their corresponding simulated counterparts. To estimate the position frequencies of each functional group in a given network, we added the frequencies of those species that belong to the group, and then, we normalised the resulting frequencies by dividing the position measure for each group by the total number of times that a group appears in any position within the same motif size class. Then, we calculated the percentile of the observed position frequencies for each group and network, just like we did with motif frequencies. To outline the general patterns of position frequencies across networks and functional groups, we fit a LMM per motif position, where the response variable was the observed position percentile per network. We used the functional group identifier as an explanatory variable and the study identifiers as a random intercept. By adding the functional group estimates to the model intercept, we assessed the average motif frequency, after controlling the variation at the study level. Finally, we visualized with the help of the package _ComplexHeatmap_ version _2.6.2_ [@gu2016] over- and under- representation of plant and pollinator functional groups on the different motif positions.

**Over- and under-represented funtional group combinations of motifs**

Finally, we studied which combinations of functional groups tend to appear together within the same motifs (up to five nodes). That is which combinations are over or under-represented. This analysis use 57 out of the 60 networks available due to computational limitations to identify all the nodes in the motifs of the three networks with the highest number of links. To do so, for each of the 53,250 possible motif combinations, we estimated the observed and the expected probability of finding those combination in empirical networks, respectively. Then, we determined whether the observed probabilities are likely to come from the expected probabilities or not. To calculate the observed probability ($p_i^O$) of functional groups within a motif $i$ (e.g., motif 3, 'bee' + 'bee' + 'selfing herbs'), we divided the number of times that $i$ appears in our set of empirical networks ($n_i^O$) by the sum of the number of times that each possible combination appeared: $p_i^O = n_i^O / \sum_{k=1}^{53,250} n_k^O$. To estimate the expected probability of a given motif combination $i$, $p_i^E$, firstly, we calculated the probability of finding a given functional group $x$ in the position $\alpha$ of $i$, $p_i(x,\alpha)$. Then, by assuming the independency of $p_i(x,\alpha)$, we computed the expected probability of the combination $i$ as the product of the probability of its pairs $(x,\alpha)$, that is, $p_i^E = \prod _{(x,\alpha )\in i} p_i(x,\alpha)$. To obtain $p_i(x,\alpha)$, we proceeded as follows. First, we calculated the number of times that the functional group $x$ appears in the position $\alpha$ in each empirical network $\eta$, denoted as $n_i^{\eta}(x,\alpha)$. Then, by controlling the variation at the network level, we assessed the average value of the absolute frequency of functional group $x$ appears in the position $\alpha$, $n_i(x,\alpha) = E\left [ n_i^{\eta}(x,\alpha) \right ]$, and estimated $p_i(x,\alpha)$ as $p_i(x,\alpha) = n_i(x,\alpha)/\sum_{k}n_i(k,\alpha)$. To obtain $n_i(x,\alpha)$, we fitted a LMM per motif position, where the response variable was the number of times that a given position was observed per network, the explanatory variable was functional group identifier, and the random intercept was given by network identifiers nested within the study identifiers.

Once we obtained $p_i^O$ and $p_i^E$, we used a simulation approach to determine whether the former is likely to come from the latter or not. This approach was preferred since the large number of possible motif combinations and the small probabilities for some of them advise against using an exact test of goodness-of-fit or a Chi-square one. Specifically, we created 1,000 random samples with repetition of possible motif combinations, where each sample contained 10 million elements and, for each combination, the probability of being selected was equal to its expected probability. From those random samples, we extracted the mean and the standard deviation of the expected probability of $i$, denoted as $E\left [ p_i^E \right ]$ and $\sigma\left [ p_i^E \right ]$, respectively, and calculated the z-scores of $p_i^O$ as $z_i^O = \left ( p_i^O - E\left [ p_i^E \right ] \right )/\sigma \left [ p_i^E \right ]$, for those motif combinations with $p_i^O > 0$. According to the usual interpretation of z-scores, combinations with $z_i^O > 1.96$ are over-represented, whereas those with $z_i^O < -1.96$ are under-represented, at the 95$\%$  confidence level. Notice that we focused on combinations with $p_i^O > 10^{-7}$ because, beyond the interest that can arouse the combinations that appear at most once, we do not have enough numerical resolution to accurately detect whether or not those combinations are under-represented (due to the limited size of our random samples).

#RESULTS

**Plant and floral visitor functional groups**

The hierarchical cluster analysis divided the dataset into five different groups with different and overlapping characteristics (**Figure S3** and **Figure S4**). The first group referred to as 'selfing herbs' consisted on herbs with hermaphroditic flowers with high levels of autonomous selfing. The second group named 'small outcrossing perennials' had small perennial species with a mixed of life forms (i.e., trees shrubs and herbs) with outcrossing hermaphroditic flowers. The third group referred to as 'self-incompatible perennials with large flowers' comprised perennial species with a mixed of life forms and large self-incompatible hermaphroditic flowers with high number of ovules. The fourth group named 'tall plants with small unisexual flowers' had the tallest species, highest proportion of shrub and tree life forms, dioecious and monoecious breeding systems, small flowers and the highest numbers of flowers and inflorescences per plant. Finally, the fifth group named 'short-lived outcrossers with long zygomorphic flowers' consisted on small perennial and short-lived herbs with long self-compatible zygomorphic flowers that were unable to self-pollinate.

In total, there were 1,126 species of floral visitor with 6,325 interactions recorded with plants. Most plants interacted with bees (2,256 interactions) and non-syrphid Diptera (1,768) followed by syrphids (845), Lepidoptera (437), Coleoptera (432) and non-bee Hymenoptera (362).

**Overall motif patterns**

Most motifs were under- and over-represented (close to the 1st and 99th percentile, respectively) in the comparison between empirical and simulated networks (**Figure 2**). Motifs 3, 5, 9, 10 and 14 were under-represented in empirical networks, that is, all were close to the 1st percentile and under the 25th percentile. Interestingly, four out of five of these motifs belonged to the largest path length classification (i.e., core-peripheral). In addition, motifs 2, 6, 7, 16 and 17 were over-represented, all over the 75th percentile and thus the closest to the 99th percentile. In contrast to the under-represented motifs, over-represented motifs belonged to the two shortest path length groups (i.e., complete and fan). The remaining motifs (i.e., 4, 8, 11, 12, 13, and 15) were between the 25th and 75th percentile. Notably, the exclusion of singletones from this analysis did not change the observed results (**Figure S1**).

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.width=4.5,fig.height=6.5, fig.cap="**Figure 2.** Comparison of motif frequencies between empirical and simulated networks grouped by average path length (plots a, b, c and d) as determined in Simmons et al. (2020). This is shown with the mean percentage of motif frequencies in empirical networks that were over the motif frequencies of the simulated ones (percentiles). This was done by network (light blue dots) and then averaged for all networks (black dots with error bars that correspond to the standard deviation)."}
library(lme4)
library(nlme)
library(tidyverse)
source("../Scripts/Scripts_Alfonso/add_study_id.R")
d <- read_csv("../Data/Csv/Motifs_frequencies_and_null_models/Motifs_frequency_percentile.csv")
d <- add_study_id(d)
#str(d)
#head(d)

#ggplot(d %>% filter(motif != 1), aes(percentil_sizeclass))+
#  geom_histogram(color="black", fill="white")+
#  facet_wrap(~motif)+
#  theme_bw()+
#  labs(x="Sizeclass percentile")
#
# Motifs tend to be over- and under-represensented in real networks

# Mean percentile and SE taking into account the study system
motif_codes <- d %>% filter(motif != 1) %>% # We remove links (motif code: 1)
  select(motif) %>% unique() %>% pull()
motif_means <- tibble(motif = motif_codes)
motif_means$mean <- NA
motif_means$SE <- NA

for(i.motif in 1:length(motif_codes)){

  m <- lmer(percentil_sizeclass ~ 1+(1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = subset(d, motif == motif_codes[i.motif]))

  motif_means$mean[i.motif] <- fixed.effects(m) %>% unname()
  motif_means$SE[i.motif] <- sqrt(diag(vcov(m)))

}

#Add categories based on path length set by Simmons 2021 et al., Functional Ecology
d$Broad_categories[d$motif==2 | d$motif==7 | d$motif==17 | d$motif==3|d$motif==4|d$motif==8] <- "Fan"

d$Broad_categories[d$motif==15 | d$motif==11] <- "Medium-weak"

d$Broad_categories[d$motif==6 | d$motif==16 |  d$motif==12] <- "Strong"

d$Broad_categories[d$motif==5 | d$motif==9 |  d$motif==14|  d$motif==13|  d$motif==10] <- "Weak"

#order the categories from short to long path length
d$Broad_categories <- factor(d$Broad_categories, levels = c("Strong", "Fan", "Medium-weak", "Weak"))

#Ffilter out two node motifs
d_1 <- d %>% filter(motif != 1)
#Merge dataframes to bl able to create facets
d_2 <- merge(d_1, motif_means, by="motif")

arr <- list('Strong' = expression(bold(paste("(a) Complete (strong, ", bar(x), " = 1.38)"))),
             'Fan' = expression(bold(paste("(b) Fan (medium-strong, ", bar(x), " = 1.48)"))), 
             'Medium-weak' = expression(bold(paste("(c) Asymmetric complete (medium-weak, ", bar(x), " = 1.60)"))), 
             "Weak" =expression(bold(paste("(d) Core-peripheral (weak, ", bar(x), " = 1.85)"))))
mylabel <- function(val) { return(lapply(val, function(x) arr[x])) }

library(ggforce) #workaround with ggforce to have facets with same sizes

p <- ggplot()+geom_point(data = d_2, aes(y=factor(motif), 100*percentil_sizeclass), shape=21,color="black",fill="azure2",size=1.75, alpha=0.3)+
  geom_errorbar(data = d_2,aes(y = factor(motif), xmin=100*(mean-SE), xmax=100*(mean+SE)), 
                width=0.8,size=0.8,color="black") +theme_bw()+
  geom_point(data = d_2,aes(y = as.factor(motif), x=100*mean), 
             size=2,color="black")+labs(y="Two to five node motifs", x = "Percentile")+
  theme( strip.background = element_blank())+
  theme(strip.text = element_text(face="bold", size=10, hjust=0),
        panel.border=element_rect(color="black",size=1.2))

p + ggforce::facet_col(vars(Broad_categories), scales = 'free', space = 'free',labeller=mylabel)

#Trying to maintain a gradient from low to high in colors

#Fig 1 Should be a dot plot (or forest plot) of the mean +- SE of the 15 motifs.
#NICE! Can we put motifs in Y axes, and % in X?
```

**Over and under-represented functional groups on motif positions**

```{r , include=FALSE}

#SAME SCRIPT AS THE PLOTTING FOR THE HEATMAP BUT I NEED IT HERE TO LOAD THE VARIABLES WITH VALUES ON RESULTS

library(lme4)
library(nlme)
library(tidyverse)
library(stringi)
source("../Scripts/Scripts_Alfonso/add_study_id.R")

#############################################
# EXTRACT MEAN VALUES
#############################################
all_position_percentiles <- read_csv("../Data/Csv/Motifs_positions_and_null_models/GF_positions_frequency_percentile.csv")

all_position_percentiles$position <- stri_extract_first_regex(all_position_percentiles$position,
                                                              "[0-9]+") %>% as.numeric()

all_position_percentiles <- add_study_id(all_position_percentiles)
#str(all_position_percentiles)
#head(all_position_percentiles)

list_plants_FG <- as.character(1:10)

plant_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(Node_FG %in% list_plants_FG, !position %in% c(1,2),
         !is.na(expected_freq_its_GF)) #ME DA ERROR
pollinator_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(!Node_FG %in% list_plants_FG, !position %in% c(1,2),
         !is.na(expected_freq_its_GF))

# Plants-------------

# Mean percentile and SE taking into account the study system
plant_codes <- plant_position_percentiles_filtered$position %>% unique()
plant_means <- plant_position_percentiles_filtered %>% select(position,Node_FG) %>% unique()
plant_means$mean <- NA
plant_means$SE <- NA

for(i.pos in 1:length(plant_codes)){

  m <- lmer(percentil_its_GF ~ Node_FG + (1|study_id),
            data = plant_position_percentiles_filtered %>%
              filter(position == plant_codes[i.pos]))

  temp <- fixed.effects(m) %>% unname()
  plant_means$mean[plant_means$position == plant_codes[i.pos]] <- c(temp[1],
                                                                    temp[1]+temp[2],
                                                                    temp[1]+temp[3],
                                                                    temp[1]+temp[4],
                                                                    temp[1]+temp[5]) ## to get real means per FG
  temp <- sqrt(diag(vcov(m)))
  plant_means$SE[plant_means$position == plant_codes[i.pos]] <- c(temp[1], 
                                                                  temp[1]+temp[2],
                                                                  temp[1]+temp[3],
                                                                  temp[1]+temp[4],
                                                                  temp[1]+temp[5])

}

# Pollinators--------

# Mean percentile and SE taking into account the study system
pollinator_codes <- pollinator_position_percentiles_filtered$position %>% unique()
pollinator_means <- pollinator_position_percentiles_filtered %>% select(position,Node_FG) %>% unique()
pollinator_means$mean <- NA
pollinator_means$SE <- NA
pollinator_means_reordered <- NULL

for(i.pos in 1:length(pollinator_codes)){

  m <- lmer(percentil_its_GF ~ Node_FG + (1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = pollinator_position_percentiles_filtered %>%
              filter(position == pollinator_codes[i.pos]),
            control = lmerControl(optimizer ="Nelder_Mead")) # I changed the optimizer because motif 8 (i.pos = 3) caused convergence problems, max gradient = 0.0022 > 0.002.

  temp <- fixed.effects(m) %>% unname()
  pollinator_means_aux <- pollinator_means[pollinator_means$position == pollinator_codes[i.pos],] %>% arrange(Node_FG)

  pollinator_means_aux$mean <- c(temp[1],
                                 temp[1]+temp[2],
                                 temp[1]+temp[3],
                                 temp[1]+temp[4],
                                 temp[1]+temp[5],
                                 temp[1]+temp[6],
                                 temp[1]+temp[7],
                                 temp[1]+temp[8],
                                 temp[1]+temp[9]) ## to get real means per FG

  temp <- sqrt(diag(vcov(m)))
  pollinator_means_aux$SE <- c(temp[1],
                               temp[1]+temp[2],
                               temp[1]+temp[3],
                               temp[1]+temp[4],
                               temp[1]+temp[5],
                               temp[1]+temp[6],
                               temp[1]+temp[7],
                               temp[1]+temp[8],
                               temp[1]+temp[9])

  pollinator_means_reordered <- bind_rows(pollinator_means_reordered,pollinator_means_aux)

}

# There are lower limits (mean - SE) smaller than 0 (see Other insects)
# We set the lower limit of those error bars to zero

pollinator_means_reordered <- pollinator_means_reordered %>% mutate(lower = mean-SE)
pollinator_means_reordered$lower[pollinator_means_reordered$lower < 0] <- 0

#############################################
# ADDING BROAD MOTIF CATEGORIES TO POSITIONS
#############################################

motifs_raw_data <- read_csv("../Data/Data_processing/Motifs_connections/motif_pattern_connections.csv")

motifs_raw_data$Broad_categories <- NA

motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(5,9,14,13,10)] <- "Core-peripherical"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(6,16,12)] <- "Complete"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(2,3,7,4,17,8)] <- "Fan"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(15,11)] <- "Asymmetric complete"

motifs_raw_data$plant <- gsub("[^0-9.-]", "", motifs_raw_data$plant)
motifs_raw_data$pollinator <- gsub("[^0-9.-]", "", motifs_raw_data$pollinator)
motifs_raw_data$plant <- as.numeric(motifs_raw_data$plant)
motifs_raw_data$pollinator <- as.numeric(motifs_raw_data$pollinator)
motifs_raw_data <- motifs_raw_data %>% filter(!is.na(Broad_categories))

fan_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Fan"] %>% unique()
core_per_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Core-peripherical"] %>% unique()
complete_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Complete"] %>% unique()
asymm_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Asymmetric complete"] %>% unique()

pollinator_means_reordered$Broad_categories <- NA

pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% fan_pollinator] <- "Fan" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% asymm_pollinator] <- "Medium-weak" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% complete_pollinator] <- "Strong" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% core_per_pollinator] <- "Weak"

fan_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Fan"] %>% unique()
core_per_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Core-peripherical"] %>% unique()
complete_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Complete"] %>% unique()
asymm_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Asymmetric complete"] %>% unique()

plant_means$Broad_categories <- NA

plant_means$Broad_categories[plant_means$position %in% fan_plant] <- "Fan" 
plant_means$Broad_categories[plant_means$position %in% asymm_plant] <- "Medium-weak" 
plant_means$Broad_categories[plant_means$position %in% complete_plant] <- "Strong" 
plant_means$Broad_categories[plant_means$position %in% core_per_plant] <- "Weak"

################################################################################
# Alternative code to HEATMAPS to create variables with values to load on text
# in a reproducible/automated way
################################################################################

# PLANTS-----

# Prepare data for heatmap
heat_plants_aux <- plant_means %>% spread(Node_FG,mean) %>% select(-SE)

heat_plants <- heat_plants_aux %>%
  group_by(position,Broad_categories) %>% 
  summarise_all(sum,na.rm = T)

# POLLINATORS-----------

filter_rare_groups <- T

if(filter_rare_groups == T){
  heat_pollinators_aux <- pollinator_means_reordered %>% filter(!Node_FG %in% c("Birds",
                                                                                "Lizards",
                                                                                "Other_insects")) %>% 
    spread(Node_FG,mean) %>% select(-SE,-lower)

  colnames(heat_pollinators_aux) <- c("position","Broad_categories","Bees",            
                                      "Coleoptera","Lepidoptera","Non-bee\nHymenoptera", 
                                      "Non-syrphids\nDiptera","Syrphids")
}else{
  heat_pollinators_aux <- pollinator_means_reordered %>% spread(Node_FG,mean) %>% select(-SE,-lower)

  colnames(heat_pollinators_aux) <- c("position","Broad_categories","Bees","Birds",            
                                      "Coleoptera","Lepidoptera","Lizards","Non-bee\nHymenoptera", 
                                      "Non-syrphids\nDiptera","Other insects","Syrphids")
}

heat_pollinators <- heat_pollinators_aux %>%
  group_by(position,Broad_categories) %>% 
  summarise_all(sum,na.rm = T)

specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))

#Check percentage of over and under represented
colMeans(heat_pollinators[,c(3:8)] < 0.25)
poll_under_25 <- colMeans(heat_pollinators[,c(3:8)] < 0.25)
poll_under_25 <- specify_decimal((poll_under_25[[1]] + poll_under_25[[2]] + poll_under_25[[3]] +
poll_under_25[[4]] + poll_under_25[[5]] + poll_under_25[[6]])/6*100,2)
#[1] 32.57576

colMeans(heat_pollinators[,c(3:8)] > 0.75)
poll_over_75 <- colMeans(heat_pollinators[,c(3:8)] > 0.75)
poll_over_75 <- specify_decimal((poll_over_75[[1]] + poll_over_75[[2]] + poll_over_75[[3]] +
    poll_over_75[[4]] + poll_over_75[[5]] + poll_over_75[[6]])/6*100,2)
#[1] 4.545455

#################################################################################
#################################################################################

colnames(heat_plants)<- c("position", "Broad_categories", "one",
                               "two", "three","four",
                               "five")
f <- heat_plants %>%
  mutate(min = pmin(1,
                    one, two,three,
                    four, five))%>%
  mutate(max = pmax(one,
                    two, three,four,
                    five))

diff_1 <- f$max-f$min

cbind(diff,diff_1)

#Check percentage of over and under represented per FG
colMeans(heat_plants[,c(3:7)] < 0.25)
plant_under_25 <- colMeans(heat_plants[,c(3:7)] < 0.25)
plant_under_25 <- specify_decimal((plant_under_25[[1]] + plant_under_25[[2]] + plant_under_25[[3]] + 
    plant_under_25[[4]] + plant_under_25[[5]])/5*100,2)
#[1] 18.18182
colMeans(heat_plants[,c(3:7)] > 0.75)
plant_over_75 <- colMeans(heat_plants[,c(3:7)] > 0.75)
plant_over_75 <- specify_decimal((plant_over_75[[1]] + plant_over_75[[2]] + plant_over_75[[3]] + 
    plant_over_75[[4]] + plant_over_75[[5]])/5*100,2)
#[1] 8.181818

#overall percentages
under <- specify_decimal((as.numeric(poll_under_25) + as.numeric(plant_under_25))/2,2)
over <- specify_decimal((as.numeric(poll_over_75) + as.numeric(plant_over_75))/2,2)

```

>>>>

---
title: "The non-random assembly of functional motifs in plant-pollinator networks worldwide"
output: pdf_document
---

__Jose B. Lanuza^1,2^__  | __Alfonso Allen-Perkins^2,3^__ | __Ignasi Bartomeus^2^__

@@

### 2022-03-08 Manuscript/Supplementary_Material/Indirect_direct_pollinators.Rmd
Edits on text

plant_means_reordered <- read_csv("../Data/Csv/plant_abs_freq_means.csv")
pollinator_means_reordered <- read_csv("../Data/Csv/pollinator_abs_freq_means.csv")
#Filter out in order to have just 6 FG'S
pollinator_means_reordered <- pollinator_means_reordered %>% filter(!Node_FG %in% c("Birds", "Lizards", "Other_insects"))

>>>>

pollinator_means_reordered <- read_csv("../Data/Csv/pollinator_abs_freq_means.csv")
#Filter out in order to have just 6 FG'S
pollinator_means_reordered <- pollinator_means_reordered %>% filter(!Node_FG %in% c("Birds", "Lizards", "Other_insects"))
